### 2020-09-23 paper.Rmd
push
push

---
title: "waRne - Putting cricket analytics in a spin"
authors: Robert(1), James (2)
output: 
  html_document:
    toc: true
    toc_float: true
    keep_md: true

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(out.width = "75%",
                      dev = "png",
                      fig.align = "center",
                      echo = FALSE)
```

```{r libraries, include=FALSE}
library(tidyverse)
library(naniar)
library(visdat)
library(patchwork)

```

## Abstract

The importance of reproducibility, and the related issue of open access to data, has received a lot of recent attention. Momentum on these issues is gathering in the sports analytics community. While cricket is the second largest commercial sport in the world, unlike other popular international sports, there has been no mechanism for the public to access comprehensive statistics on players and teams. Expert commentary currently relies heavily on data that isn't made readily accessible and this produces an unnecessary barrier for the development of an inclusive sports analytics community.

## Introduction

## What is waRne

We need to think about what it is waRne should be?

1.  provide the csv we used for datajam and then people just add games to it

2.  people can not have the csv but just download random tournaments

## Why cricket needs reproducibility

maybe include stuff around cricviz interesting tweets but fans can recreate or investigate further.

## Cool cricket examples

-   Dhoni spin/pace/yorkr

    -   shows off joins to player information and extracting information from play by play

-   espncricinfo articles but instead of ipl do same tables for big bash

-   recreate stats from espncricinfo like this [page](https://www.espncricinfo.com/series/19297/statistics/1187665/new-zealand-vs-england-1st-t20i-england-in-new-zealand-2019-20
)

| Things to do                 | person                                                                   |
|------------------------------|--------------------------------------------------------------------------|
| clean scrapers/csv for waRne | James                                                                    |
| data dictionary              | Robert                                                                   |
| examples                     | Robert - start (assume same csv being used) - cool cricket example stuff |

: things to do