### 2020-09-24 paper.Rmd
beaten by con artists
proof?

---
title: "waRne - Putting cricket analytics in a spin"
authors: Robert(1), James (2)
output: 
  html_document:
    toc: true
    toc_float: true
    keep_md: true

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(out.width = "75%",
                      dev = "png",
                      fig.align = "center",
                      echo = FALSE)
```

```{r libraries, include=FALSE}
library(tidyverse)
library(naniar)
library(visdat)
library(patchwork)

```

## Abstract

The importance of reproducibility, and the related issue of open access to data, has received a lot of recent attention. Momentum on these issues is gathering in the sports analytics community. While cricket is the second largest commercial sport in the world, unlike other popular international sports, there has been no mechanism for the public to access comprehensive statistics on players and teams. Expert commentary currently relies heavily on data that isn't made readily accessible and this produces an unnecessary barrier for the development of an inclusive sports analytics community.

## Introduction

## What is waRne

We need to think about what it is waRne should be?

1.  provide the csv we used for datajam and then people just add games to it

2.  people can not have the csv but just download random tournaments

## Why cricket needs reproducibility

maybe include stuff around cricviz interesting tweets but fans can recreate or investigate further.

James and I entered a reproducibility competition and the winning entry provided a statistic sort of a gateway into the importance of their ''project'' was was specifically in 2019 Australian Big Bash T20 season. The statistic given was in the 2019 big bash season teams that batted first won only $43$%$ of games. Thankfully we can try to reproduce that statistic now through the use of waRne

Wikipedia tells us that there were the following:

In the 2018-19 Big Bash League there were 59 matches played. Of which we had 3 DLS results.

| Games                             | Result                        |
|-----------------------------------|-------------------------------|
| 21 DEC 2018 - THUNDER VS STARS    | THUNDER WON BY 15 RUNS (D/L)  |
| 2 FEB 2019 - THUNDER VS SIXERS    | SIXERS WON BY 9 WICKETS (D/L) |
| 8 JAN 2019 HEAT VS THUNDER        | HEAT BY 15 RUNS (D/L)         |
| 17th January 2019 Thunder vs Heat | no result                     |

: Games under consideration in 2018-19 Big Bash

```{r, investigation}
df<-read.csv("data/combined_2020_09_11.csv")

blah<-df%>%
	dplyr::select(duckworth_lewis,year,match_id, tournament_name, tournament_type, team, inning, binary_result)%>%
	dplyr::filter(tournament_type%in% c("Twenty20 Internationals", "Twenty20"  ))%>%
	dplyr::filter(tournament_name%in% c("Twenty20 Big Bash","Big Bash League" ))%>%
									dplyr::filter(inning==1)%>%
									dplyr::group_by()%>%
									distinct()
# dataset doesn't contain duckworth lewis results. 
table(blah$year,blah$binary_result)
````
```

So here we can see that we in the big bash season 2018-19 we have a current winning percentage batting first of 22/(22+33) =0.4 but this doesn't include the duckworth lewis games or the washed out completely game.

If we were to add in the duckworth lewis games we get 24/(24+34) which is equal to 0.414. Which isn't the result.

We can also clearly see that in the 20192020 big bash season the team batting first wins 39/(39+20) or 0.66 so what could be going on? Did Quantium mean only games played in the actual year 2019?

In that case from the 2018-2019 season we would remove the following games

| Date                 | Game                    | Result                   | Winning Team |
|----------------------|-------------------------|--------------------------|--------------|
| 19 December 2018     | Heat Vs Strickers       | Strickers 5 Wickets      | Chased       |
| 20th December 2018   | Scorchers Vs Renegades  | Renegades by 4 wickets   | Chased       |
| 21 December 2018     | Thunder vs Stars        | Thunder by 15 runs (D/L) | Batted First |
| 22 December 2018     | Sixers vs Scorchers     | Sixers by 17 runs        | Batted First |
| 22 December          | Hurricanes vs Heat      | Hurricanes by 15 runs    | Batted First |
| 23 December 2018     | Strikers vs Renegades   | Renegades by 5 wickets   | Chased       |
| 24 December 2018     | Stars vs Hurricanes     | Hurricanes by 6 wickets  | Chased       |
| 24th December 2018   | Thunder vs sixers       | Thunder by 21 runs       | Batted First |
| 26th December 2018   | Strickers vs scorchers  | Scorchers by 7 wickets   | Chased       |
| 27th December        | Sixers vs Stars         | stars by 5 wickets       | Chased       |
| 28th December        | thunder vs hurricanes   | hurricanes by 7 wickets  | Chased       |
| 29th december        | sixers vs renegades     | sixers by 33 runs        | Batted First |
| 30th decemeber 2018  | scorchers vs hurricanes | hurricanes by 6 wickets  | Chased       |
| 31st decemember 2018 | strickers vs thunder    | strickers by 20 runs     | Batted First |

: Games in 2018 of the 2018-19 Big Bash Season

i.e. we remove 14 games here of which 6 teams won batting first 16/(22+33-14)

Then from here we have to add on games in the 2019-2020 big bash season that were played in 2019

+--------------------+-------------------------+---------------------------------------+--------------------------+
| Date               | Teams                   | Result                                | Winning Team             |
+====================+=========================+=======================================+==========================+
| 17 December        | Thunder vs Heat         | Thunder by 29 runs                    | Batted First             |
+--------------------+-------------------------+---------------------------------------+--------------------------+
| 18th December      | Scorchers vs Sixers     | Sixers by 8 wickets                   | Chased                   |
+--------------------+-------------------------+---------------------------------------+--------------------------+
| 19th december      | Renegades vs thunder    | thunder by 6 wickets                  | chased                   |
+--------------------+-------------------------+---------------------------------------+--------------------------+
| 20th decemeber     | hurricanes vs sixers    | hurricanes by 25 runs                 | batted first             |
+--------------------+-------------------------+---------------------------------------+--------------------------+
| 20th decemember    | stars vs heat           | stars by 22 runs                      | batting first            |
+--------------------+-------------------------+---------------------------------------+--------------------------+
| 21st decemember    | strickers vs thunder    | no result                             | no result                |
+--------------------+-------------------------+---------------------------------------+--------------------------+
| 21st december      | scorchers vs renegades  | scorchers by 11 runs                  | batting first            |
+--------------------+-------------------------+---------------------------------------+--------------------------+
| 22 deceember       | stars vs hurricanes     | stars by 52 runs                      | batting first            |
+--------------------+-------------------------+---------------------------------------+--------------------------+
| 22 decemember      | heat vs sixers          | heat by 48 runs                       | batting first            |
+--------------------+-------------------------+---------------------------------------+--------------------------+
| 23 dec             | strickers vs scorchers  | strickers by 15 (DLS)                 | batting first            |
+--------------------+-------------------------+---------------------------------------+--------------------------+
| 24th dec           | renegades vs hurricanes | hurricanes by 7 wickets               | chased                   |
+--------------------+-------------------------+---------------------------------------+--------------------------+
| 26 decemember      | sixers vs scorchers     | sixers by 48 runs                     | batting first            |
+--------------------+-------------------------+---------------------------------------+--------------------------+
| 27 dec             | strickers vs stars      | strikers by 5 runs                    | batting first            |
+--------------------+-------------------------+---------------------------------------+--------------------------+
| 28th december      | thunder vs sixers       | sixers won super over (batting first) | sixers by superover      |
|                    |                         |                                       |                          |
|                    |                         |                                       | or sixers batting second |
+--------------------+-------------------------+---------------------------------------+--------------------------+
| 29th dec           | strikers vs renegades   | strikers by 18 runs                   | batting first            |
+--------------------+-------------------------+---------------------------------------+--------------------------+
| 30th dec           | hurricanes vs stars     | stars by 4 runs (DLS)                 | batting first            |
+--------------------+-------------------------+---------------------------------------+--------------------------+
| 31st december 2019 | thunder vs strikers     | thunder by 3 runs                     | batting first            |
+--------------------+-------------------------+---------------------------------------+--------------------------+
|                    |                         |                                       |                          |
+--------------------+-------------------------+---------------------------------------+--------------------------+

: Games played in 2019 of the 2019-20 Big Bash

## Cool cricket examples

-   Dhoni spin/pace/yorkr

    -   shows off joins to player information and extracting information from play by play

-   espncricinfo articles but instead of ipl do same tables for big bash

-   recreate stats from espncricinfo like this [page](https://www.espncricinfo.com/series/19297/statistics/1187665/new-zealand-vs-england-1st-t20i-england-in-new-zealand-2019-20)

+------------------------------+--------------------------------------------------------------------------+
| Things to do                 | person                                                                   |
+==============================+==========================================================================+
| clean scrapers/csv for waRne | James                                                                    |
+------------------------------+--------------------------------------------------------------------------+
| data dictionary              | Robert                                                                   |
+------------------------------+--------------------------------------------------------------------------+
| examples                     | Robert - start (assume same csv being used) - cool cricket example stuff |
+------------------------------+--------------------------------------------------------------------------+

: things to do

## How would you recommend evaluating the selection of players to reduce the required run rate? Is there a tool that would could be used to let coaches or selectors do this themselves or is it would it be more appropriate to have an analyst do this prior?

##